Illuminating a Dark Lens : A Type Ia Supernova \\ Magnified by Galaxy Cluster Abell 2744

### 2015-04-04 snTomas.tex
finished sn06bt comparison section

\change{
One might suppose that a bias could be introduced if we have adopted
incorrect values for the color correction parameter $\beta$ or
extinction law $R_V$ in the SALT2 and MLCS2k2 fits, respectively.  The
appropriate value to use for this color correction and how it affects
inferences about the intrinsic scatter in Type Ia SN luminosities is a
complex question that is beyond the scope of this work \citep[see
e.g.][]{Marriner:2011,Chotard:2011,Kessler:2013,Scolnic:2014a}.
However, we can already rule this out as a solution for the
magnification discrepancy.  Our error on the \tomas\ distance modulus
already includes an uncertainty in the extinction law and a related
error to account for the intrinsic luminosity scatter. These are well
vetted paraemeters, based on observations of $\sim500$ SNe extending
to $z\sim1.5$ \citep{Sullivan:2011}.  Furthermore, changing the color
correction applied to SN \tomas\ would requre the same adjustment to
be applied to the unlensed SNe at similar redshift that make up our
comparison sample, unless one proposes without evidence that
SN \tomas\ is uniquely affected by a peculiar type of dust.  
}

@@

\textcolor{red}{TODO : discuss 2006bt-like misclassification}

\change{
In Section~\ref{sec:LightCurveFitting} we found that SN \tomas\ is on
the blue end of the normal range of Type Ia SN colors.  With the SALT2
fitter we measured a color parameter $c=-0.115\pm0.025$, and with
MLCS2k2 we found the host galaxy dust extinction to be $A_{\rm
V}=0.014\pm0.037$ magnitudes.  These colors are tightly constrained,
as we are fitting to photometry that covers a rest-frame wavelength
range from $\sim3500-7000$\AA and extends to $\sim$30 days past
maximum brightness.  
}

\textcolor{red}{TODO : discuss 2006bt-like colors}

@@

\begin{enumerate}
\item The close proximity of a cluster member galaxy is leading to a
biased magnification, because that galaxy's M/L is atypical.
\item The models are limited by a scarcity of strong lensing constraints
\item The cluster mass profile exhibits a sudden change outside of the core
strong-lensing region, where we lack multiply-imaged background
sources to constrain it. 
\item \change{The redshift assigned to SN \tomas\ is in error.}
\item \change{The color corrections for SN \tomas\ underestimate the foreground dust.}
\item \change{SN \tomas\ is not a normal Type Ia SN.}
\end{enumerate}

@@

\change{
The last two possible explanations presume an error in the
interpretation of the available SN data.  We reject the possibility
that a redshift error is the primary cause of the discrepancy, as the
redshift evidence is derived principally from the SN itself and not
from the host galaxy. A value of $z\sim1.3$ is well supported by both
the spectroscopic and photometric SN data, and changing the redshift
within the constraints of these complementary observations only serves
to increase the tension between the observed and predicted
magnifications.
}

\change{
Finally, we have considered whether the SN could be
mis-classified. Once again the combination of spectroscopic and
photometric evidence strongly supports our classification of
SN \tomas\ as a normal Type Ia SN.  The most plausible
mis-classification would be that the object is a peculiar Type Ia of
the SN 2006bt-like sub-class.  We find this possibility to be 
 \textcolor{red}{TODO: summarize the 06bt misclassification discussion}
}

>>>>

\change{
In Section~\ref{sec:LightCurveFitting} we found that SN \tomas\ is on
the blue end of the normal range of Type Ia SN colors.  With the SALT2
fitter we measured a color parameter $c=-0.115\pm0.025$, and with
MLCS2k2 we found the host galaxy dust extinction to be $A_{\rm
V}=0.014\pm0.037$ magnitudes.  These colors are tightly constrained,
as we are fitting to photometry that covers a rest-frame wavelength
range from $\sim3500-7000$\AA and extends to $\sim$30 days past
maximum brightness.  This leaves little room for the luminosity
measurement to be biased by dust, as the dimming of \tomas\ would also
necessarily be accompanied by some degree of reddening. 
}

\change{
Nevertheless, one might suppose that a bias could be introduced if we
have adopted incorrect values for the color correction parameter
$\beta$ or extinction law $R_V$ in the SALT2 and MLCS2k2 fits,
respectively.  The appropriate value to use for this color correction
and how it affects inferences about the intrinsic scatter in Type Ia
SN luminosities is a complex question that is beyond the scope of this
work \citep[see
e.g.][]{Marriner:2011,Chotard:2011,Kessler:2013,Scolnic:2014a}.
However, we can already rule this out as a solution for the
magnification discrepancy.  Our error on the \tomas\ distance modulus
already includes an uncertainty in the extinction law and a related
error to account for the intrinsic luminosity scatter. These are well
vetted paraemeters, based on observations of $\sim500$ SNe extending
to $z\sim1.5$ \citep{Sullivan:2011}.  Furthermore, there is no reason
to propose that SN \tomas\ is uniquely affected by a peculiar type of
dust.  Thus, any change in the color correction applied to SN \tomas\
would require the same adjustment to be applied to the unlensed SNe at
similar redshift that make up our comparison sample, largely negating
the effect on the inferred magnification.  }

@@

\change{
Is it possible that \tomas\ is an example of a peculiar stellar
explosion that does not follow the relationship between light curve
shape and luminosity observed in normal Type Ia SNe?  In
Sections~\ref{sec:SNID} and \ref{sec:PhotometricClassification} we
classified SN Tomas as a normal Type Ia SN based on both the
spectroscopic and photometric evidence.  This rules out the
possibility that \tomas\ belongs to a different class of normal SN
explosions (Type Ib, Ic, or II). In
Section \ref{sec:LightCurveFitting} we fit the \tomas\ light curve to
determine a luminosity distance and found that the light curve shape
and color are consistent with a Type Ia SN of average light curve
width, with minimal dust extinction.  This excludes the possibility
that \tomas\ is one of the sub-class of ``fast declining'' Type Ia SNe
like SN 1991bg.
}

\change{
One remaining possibility is that \tomas\ could be part of a rare
sub-category of peculiar Type Ia SNe that masquerade as their normal
cousins, epitomized by the prototype SN 2006bt \citep{Foley:2010}.
These objects appear to have a normal Type Ia light curve shape,
except for the absence of a secondary maximum or ``shoulder'' in
near-IR bands.  The reddest filter available for the \tomas\ light
curve is F160W, which has an effective rest-frame wavelength of
6623\AA\ at $z=1.31$, making it close to the rest-frame {\it r} band.
Although the near-IR shoulder is more prominent in the rest-frame {\it
i} band, we can see in Figures~\ref{fig:LightCurveFits} that the F160W
light curve may suggest a weak or delayed near-IR shoulder.  The
second-to-last observation in F160W is $\sim1\sigma$ lower than
predicted by the best-fit SALT2 and MLCS2k2 models. This is tenuous
evidence, but it would be consistent with a 2006bt-like light curve.
}

\change{
In this case, however, the color of SN \tomas\ can rule out a
classification as a 2006bt-like object.  All members of this peculiar
sub-class exhibit very red colors across all optical and near-IR
bands, consistent with a very cool photosphere \citep{Foley:2010}.
Figure~\ref{fig:colorcomparison} shows that \tomas\ is bluer than the
SN 2006bt prototype by at least 0.25 magnitudes in every color. The
observed \tomas\ colors are fully consistent with the a normal, blue
Type Ia SN (the SALT2 model is shown).
}

@@

\begin{enumerate}
\item The close proximity of a cluster member galaxy is leading to a
biased magnification, because that galaxy's M/L is atypical.
\item The models are limited by a scarcity of strong lensing constraints
\item The cluster mass profile exhibits a sudden change outside of the core
strong-lensing region, where we lack multiply-imaged background
sources to constrain it. 
\item \change{The redshift assigned to SN \tomas\ is in error.}
\item \change{SN \tomas\ is dimmed by foreground dust.}
\item \change{SN \tomas\ is not a normal Type Ia SN.}
\end{enumerate}

@@

\change{
The last three possible explanations presume an error in the
interpretation of the available SN data.  We reject the possibility
that a redshift error is the primary cause of the discrepancy, as the
redshift evidence is derived principally from the SN itself and not
from the host galaxy. A value of $z\sim1.3$ is well supported by both
the spectroscopic and photometric SN data, and changing the redshift
within the constraints of these complementary observations only serves
to increase the tension between the observed and predicted
magnifications.  We also find it implausible that there is sufficient
dust in the cluster or elsewhere along the line of sight to account
for the magnification discrepancy.  A dust-induced bias is also
disfavored by the very blue color of SN \tomas.  }

\change{
Finally, we have considered whether the SN could be
mis-classified. Once again the combination of spectroscopic and
photometric evidence strongly supports our classification of
SN \tomas\ as a normal Type Ia SN.  The most plausible
mis-classification would be that the object is a peculiar Type Ia of
the SN 2006bt-like sub-class.  Although the light curve shape could
allow this possibility, the blue color of \tomas\ can once again
reject this alternative. 
}
